Significant Genes and Enriched Pathways

### 2020-11-24 StanAnalysis_updated.Rmd
Edits to figures

fBG <- function(x, a, m1, s1, m2, s2) a*dnorm(x,m1,s1) + (1-a)*dnorm(x,m2,s2) # bimodal gaussian
post_prior_histogram <- function(postvals,mask,priorfun,bw,xl,titlestr) {
  # stacked histogram colored by gene-associated sites
  # and prior curve overlaid
  plot_df <- data.frame(val=postvals, gene_assoc=mask)
  p1 <- ggplot(plot_df, aes(x=val,fill=gene_assoc)) +
    #geom_histogram(aes(y=..count../length(mask)), binwidth=bw, position="dodge") +
    #stat_function(fun=function(x) bw*priorfun(x),geom="line",color="black") +
    geom_histogram(aes(y=..density..), binwidth=bw, position="dodge") +
    stat_function(fun=function(x) priorfun(x),geom="line",color="black") +
    ggtitle(titlestr) + xlim(xl) + theme_classic() + theme(legend.position = "none") + xlab(titlestr) + ylab("Freq.")
  p1
}

# plot mu,betaT,sigmaE,sigmaP,sigmaPT,sigmaT with priors overlaid
pM <- post_prior_histogram(mu_full$p50, genemask, function(x) 1.3*fBG(x,0.3557,-3.0675, 1.3027, 1.1290, 2.4110), 0.25, c(-5,5), "Mu")
pB <- post_prior_histogram(betaT_full$p50, genemask, function(x) 1.5*dcauchy(x,-0.04051, 0.5799), 0.25, c(-5,5), "BetaT")
pE <- post_prior_histogram(sigmaE_full$p50, genemask, function(x) 3.3*dgamma(x,2,2), 0.05, c(0,2), "SigmaE")
pP <- post_prior_histogram(sigmaP_full$p50, genemask, function(x) 1.52*dgamma(x, 2.7693, 7.0029), 0.05, c(0,2), "SigmaP")
pPT <- post_prior_histogram(sigmaPT_full$p50, genemask, function(x) 1.6*dgamma(x, 2.1457, 2.4819), 0.05, c(0,2), "SigmaPT")
pT <- post_prior_histogram(sigmaT_full$p50, genemask, function(x) 0.97*dgamma(x, 3.3643, 10.430), 0.05, c(0,2), "SigmaT")
# log sigmas
pLP <- post_prior_histogram(log(sigmaP_full$p50), genemask, function(x) 0.34*dgamma(exp(x), 2.7693, 7.0029), 0.25, c(-5,5), "log(SigmaP)")
pLPT <- post_prior_histogram(log(sigmaPT_full$p50), genemask, function(x) 0.42*dgamma(exp(x), 2.1457, 2.4819), 0.25, c(-5,5), "log(SigmaPT)")
pLT <- post_prior_histogram(log(sigmaT_full$p50), genemask, function(x) 0.22*dgamma(exp(x), 3.3643, 10.430), 0.25, c(-5,5), "log(SigmaT)") + theme(legend.position = c(0.8, 0.6))

# arrange each parameter as a subplot # ,labels=c("M","B","E","P","PT","T","lP","lPT","lT")
p_all <- ggarrange(pM,pB,pE,pP,pPT,pT,pLP,pLPT,pLT,hjust=0,nrow=3,ncol=3)
print(p_all)
# ggsave("posteriors_geneassoc.svg",p_all)
rm(pM,pB,pE,pP,pPT,pT,pLP,pLPT,pLT)
```

@@

# plot the two conservations score prob sums
pP1 <- post_histogram(CpGscore_full$median,genemask,0.05,c(-3,3),expression("Median log("~sigma[P]/sigma[T]~")")) + scale_x_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3))
prob1_proxy = log(sigmaP_full$p50/sigmaT_full$p50)
pP2 <- post_histogram(prob1_proxy,genemask,0.05,c(-3,3),expression(" log( median" ~sigma[P]/~"median"~sigma[T]~")"))
# regularized version
L2norm = sqrt(sigmaP_full$p50^2 + sigmaT_full$p50^2)
prob1_reg = log(sigmaP_full$p50/sigmaT_full$p50) - 1/L2norm
pP3 <- post_histogram(prob1_reg,genemask,0.1,c(-10,5),"Regularized")
p_probs <- ggarrange(pP1,pP2,pP3,nrow=1,ncol=3)
print(p_probs)
```

@@

```{r load_prep, echo=FALSE, include=TRUE, message=FALSE, eval=!params$prep}
cat("Param:prep=FALSE -> load saved data (BOTH)...")
gened_data <- read_csv(here(prepdatafolder, "gened_data.csv"), col_types = list("CHR" = col_character()))
prepped_data <- read_csv(here(prepdatafolder, "prepped_data.csv"))
permed_data <- read_csv(here(prepdatafolder, "permed_data.csv"))
promoted_data <- read_csv(here(prepdatafolder, "promoted_data.csv"))
cat("done")
```

@@

# plot CpG island relation score distributions
prepped_data %>% select(Name, Relation_to_UCSC_CpG_Island, score_median) %>%
  dplyr::rename(ir = Relation_to_UCSC_CpG_Island, score=score_median) %>%
  mutate(ir = case_when(ir=="" ~ "Sea",
                        str_detect(ir, "^S") ~ str_replace(ir, "^S_", ""),
                        str_detect(ir, "^N") ~ str_replace(ir, "^N_", ""),
                        TRUE ~ ir)) %>%
  mutate(ir = factor(ir, levels = c("Island", "Shore", "Shelf", "Sea"))) %>%
  #ggplot() + geom_density(aes(x=score,y=..scaled..,color=ir)) + theme_classic() + scale_x_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3)) + ylab("Normalized Density")
  ggplot() + geom_violin(aes(x=ir,y=score,fill=ir),draw_quantiles = 0.5) + theme_classic() + scale_y_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3))
# ggsave("CpGscore_islands.svg")

# plot functional region score distributions
temp <- prepped_data %>% select(Name, UCSC_RefGene_Group, score_median) %>%
  separate_rows(UCSC_RefGene_Group, sep=";") %>%
  distinct() %>%
  dplyr::rename(fr = UCSC_RefGene_Group, score=score_median) %>%
  mutate(fr = case_when(fr=="" ~ "NA",
                        str_detect(fr, "5URT") ~ "5'UTR",
                        str_detect(fr, "3UTR") ~ "3'UTR",
                        TRUE ~ fr)) %>%
  distinct()
keepregions = c("TSS200","5'UTR","Body","3'UTR")
# temp %>%
#   filter(fr %in% keepregions) %>%
#   #mutate(fr = factor(fr, levels = c("Island", "Shore", "Shelf", "Sea"))) %>%
#   ggplot() + geom_density(aes(x=score,color=fr)) + theme_classic() + scale_x_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3)) + ylab("Normalized Density")
temp %>%
  mutate(fr = factor(fr, levels = keepregions)) %>%
  filter(fr %in% keepregions) %>%
  ggplot() + geom_violin(aes(x=fr,y=score,fill=fr),draw_quantiles = 0.5) + theme_classic() + scale_y_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3))
# ggsave("CpGscore_regions_violin.svg")
rm(temp)
```

@@

# plot: n_sites, mu, betaT, sigmaP, sigmaPT, sigmaT
n_cutlo = 5
n_cuthi = quantile(gene_summary$n, 0.99)
pN <- post_histogram(gene_summary$n,1,c(0,200),paste0("N sites (max = ",max(gene_summary$n),")")) + geom_vline(xintercept=n_cutlo-0.5,linetype = "dashed",color="red") + annotate(geom="text",x=80,y=0.03,size=2,label=paste0("# genes <",n_cutlo," sites:\n", sum(gene_summary$n<n_cutlo)," / ",nrow(gene_summary))) +
geom_vline(xintercept=n_cuthi,linetype = "dashed",color="red") +annotate(geom="text",x=120,y=0.013,size=2,label=sprintf("# genes >%0.1f sites:\n %i / %i (99th pctile)", n_cuthi,sum(gene_summary$n>n_cuthi),nrow(gene_summary)))
gene_summary <- gene_summary %>% filter(n>=n_cutlo) %>% filter(n<n_cuthi)
pM <- post_histogram(gene_summary$mu,0.1,c(-5,5),"Mu")
pB <- post_histogram(gene_summary$betaT,0.1,c(-5,5),"BetaT")
pP <- post_histogram(gene_summary$sigmaP,0.05,c(0,2),"SigmaP")
pPT <- post_histogram(gene_summary$sigmaPT,0.05,c(0,2),"SigmaPT")
pT <- post_histogram(gene_summary$sigmaT,0.05,c(0,2),"SigmaT")
pP1 <- post_histogram(gene_summary$score_median,0.05,c(-3,3),"Score_median")
pP2 <- post_histogram(gene_summary$score_median2,0.05,c(-3,3),"Score_median2")
pP3 <- post_histogram(gene_summary$score_reg,0.05,c(-3,3),"Score_reg")

#pP1 <- post_histogram(gene_summary$score_prob,0.01,c(0,1),"Score_prob")
#pP2 <- post_histogram(gene_summary$score_mean,0.05,c(-3,3),"Score_mean")

# arrange each parameter as a subplot
p_all <- ggarrange(pN,pM,pB,pP,pPT,pT,pP1,pP2,pP3,hjust=0,nrow=3,ncol=3)
print(p_all)
rm(pN,pM,pB,pP,pPT,pT,pP1,pP2,pP3)

@@

#pP1 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_prob), alpha=0.01) + ggtitle("Score_prob vs N-sites")
#pP2 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_mean), alpha=0.01) + ggtitle("Score_mean vs N-sites")
pP1 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_median), alpha=0.01) + ggtitle("Score_median vs N-sites")
pP2 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_median2), alpha=0.01) + ggtitle("Score_median2 vs N-sites")
pP3 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_reg), alpha=0.01) + ggtitle("Score_reg vs N-sites")

# arrange each parameter as a subplot
p_all <- ggarrange(pM,pB,pE,pP,pPT,pT,pP1,pP2,pP3,hjust=0,nrow=3,ncol=3)
print(p_all)
rm(pM,pB,pE,pP,pPT,pT,pP1,pP2,pP3)
```

@@

distance_length <- distanced_data %>%
  filter(!is.na(UCSC_RefGene_Name)) %>%
  select(CHR, UCSC_RefGene_Name, cpgdist, c(sigmaP:sigmaT,score_median:score_reg)) %>%
  mutate(location_bin = cut(cpgdist, c(seq(-1500, 20000, by = 100)), include.lowest = TRUE)) %>%
  group_by(location_bin) %>%
  summarise(mean_sigmaP = mean(sigmaP), mean_sigmaPT = mean(sigmaPT), mean_sigmaT = mean(sigmaT),  mean_score_median = mean(score_median), mean_score_median2 = mean(score_median2), n_obs = n(), .groups='drop') %>%
  mutate(location_bin = as.numeric(str_extract(location_bin, "(?<=,)(.*)(?=])")))

@@

# again for conservation scores
distance_length %>%
  drop_na() %>%
  pivot_longer(mean_score_median:mean_score_median2) %>% 
  mutate(name = factor(name, c("mean_score_median","mean_score_median2"), c("Score_median","Score_median2"))) %>%
  sample_frac() %>%
  ggplot(aes(location_bin, value, color = name, fill = name)) +
  geom_point(key_glyph = draw_key_rect) +
  ggforce::facet_zoom(location_bin < 5000) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(title = NULL,
       fill = NULL,
       color = NULL,
       x = "Distance from gene start site (binned to 100 bp)",
       y = "Average Conservation Score", 
       caption = "Figure 1: Average conservation scores of single CpGs as a function of position relative to gene.") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
cat("\n")

@@

cat("***")
permed_data %>%
  pivot_longer(dplyr::matches("1|3")) %>%
  separate(name, into = c("type", "method"), sep = "_v") %>%
  mutate(method = factor(method, c("1", "3"), c("Naive bootstrap", "Adjusted bootstrap")),
         type = factor(type, c("score_median","score_median2","score_reg"), c("Score_median","Score_median2","Score_reg"))) %>%
  drop_na() %>%
  ggplot(aes(value, fill = type)) +
  geom_histogram(bins = 50, color = "grey30") +
  facet_grid(method~type, scales = "free") +
  theme_minimal() +
  labs(x = "p-value",
       y = "Number of genes",
       title = NULL,
       caption = "Figure 2B: Distribution of boot-strapping p-values for genes as a function of conservation score and bootstrap method.")  +
  theme(legend.position = "none") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
```

>>>>

fBG <- function(x, a, m1, s1, m2, s2) a*dnorm(x,m1,s1) + (1-a)*dnorm(x,m2,s2) # bimodal gaussian
post_prior_histogram <- function(postvals,mask,priorfun,bw,xl,titlestr) {
  # stacked histogram colored by gene-associated sites
  # and prior curve overlaid
  plot_df <- data.frame(val=postvals, gene_assoc=mask)
  p1 <- ggplot(plot_df, aes(x=val)) + #,fill=gene_assoc
    #geom_histogram(aes(y=..count../length(mask)), binwidth=bw, position="dodge") +
    #stat_function(fun=function(x) bw*priorfun(x),geom="line",color="black") +
    geom_histogram(aes(y=..density..), binwidth=bw, position="dodge", fill="lightgray") +
    stat_function(fun=function(x) priorfun(x), geom="line", color="black") +
    ggtitle(titlestr) + xlim(xl) + theme_classic() + theme(legend.position = "none") + xlab(titlestr) + ylab("Freq.")
  p1
}

# plot mu,betaT,sigmaE,sigmaP,sigmaPT,sigmaT with priors overlaid
pM <- post_prior_histogram(mu_full$p50, genemask, function(x) 1.3*fBG(x,0.3557,-3.0675, 1.3027, 1.1290, 2.4110), 0.25, c(-5,5), "Mu") + theme(aspect.ratio = 1)
pB <- post_prior_histogram(betaT_full$p50, genemask, function(x) 1.5*dcauchy(x,-0.04051, 0.5799), 0.25, c(-5,5), "BetaT") + theme(aspect.ratio = 1)
pE <- post_prior_histogram(sigmaE_full$p50, genemask, function(x) 3.3*dgamma(x,2,2), 0.05, c(0,2), "SigmaE") + theme(aspect.ratio = 1)
pP <- post_prior_histogram(sigmaP_full$p50, genemask, function(x) 1.52*dgamma(x, 2.7693, 7.0029), 0.05, c(0,2), "SigmaP") + theme(aspect.ratio = 1)
pPT <- post_prior_histogram(sigmaPT_full$p50, genemask, function(x) 1.6*dgamma(x, 2.1457, 2.4819), 0.05, c(0,2), "SigmaPT") + theme(aspect.ratio = 1)
pT <- post_prior_histogram(sigmaT_full$p50, genemask, function(x) 0.97*dgamma(x, 3.3643, 10.430), 0.05, c(0,2), "SigmaT") + theme(aspect.ratio = 1)
# log sigmas
#pLP <- post_prior_histogram(log(sigmaP_full$p50), genemask, function(x) 0.34*dgamma(exp(x), 2.7693, 7.0029), 0.25, c(-5,5), "log(SigmaP)")
#pLPT <- post_prior_histogram(log(sigmaPT_full$p50), genemask, function(x) 0.42*dgamma(exp(x), 2.1457, 2.4819), 0.25, c(-5,5), "log(SigmaPT)")
#pLT <- post_prior_histogram(log(sigmaT_full$p50), genemask, function(x) 0.22*dgamma(exp(x), 3.3643, 10.430), 0.25, c(-5,5), "log(SigmaT)") + theme(legend.position = c(0.8, 0.6))

# arrange each parameter as a subplot # ,labels=c("M","B","E","P","PT","T","lP","lPT","lT")
p_all <- ggarrange(pM,pB,pE,pP,pPT,pT,hjust=0,nrow=2,ncol=3) #,pLP,pLPT,pLT
print(p_all)
# ggsave("posteriors_geneassoc.svg",p_all)
rm(pM,pB,pE,pP,pPT,pT)
```

@@

# plot the two conservations score prob sums
pP1 <- post_histogram(CpGscore_full$median,genemask,0.1,c(-3,3),expression("Median log("~sigma[P]/sigma[T]~")")) + scale_x_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3))
# ggsave("CpGscore_geneassoc.svg")
# prob1_proxy = log(sigmaP_full$p50/sigmaT_full$p50)
# pP2 <- post_histogram(prob1_proxy,genemask,0.05,c(-3,3),expression(" log( median" ~sigma[P]/~"median"~sigma[T]~")"))
# # regularized version
# L2norm = sqrt(sigmaP_full$p50^2 + sigmaT_full$p50^2)
# prob1_reg = log(sigmaP_full$p50/sigmaT_full$p50) - 1/L2norm
# pP3 <- post_histogram(prob1_reg,genemask,0.1,c(-10,5),"Regularized")
# p_probs <- ggarrange(pP1,pP2,pP3,nrow=1,ncol=3)
# print(p_probs)
```

@@

```{r load_prep, echo=FALSE, include=TRUE, message=FALSE, eval=!params$prep}
cat("Param:prep=FALSE -> load saved data (BOTH)...")
gened_data <- read_csv(here(prepdatafolder, "gened_data.csv"), col_types = list("CHR" = col_character()))
prepped_data <- read_csv(here(prepdatafolder, "prepped_data.csv"))
permed_data <- read_csv(here(prepdatafolder, "permed_data.csv"))
cat("done")
```

@@

# plot CpG island relation score distributions
prepped_data %>% select(Name, Relation_to_UCSC_CpG_Island, score_median) %>%
  dplyr::rename(ir = Relation_to_UCSC_CpG_Island, score=score_median) %>%
  mutate(ir = case_when(is.na(ir) ~ "Sea",
                        str_detect(ir, "^S") ~ str_replace(ir, "^S_", ""),
                        str_detect(ir, "^N") ~ str_replace(ir, "^N_", ""),
                        TRUE ~ ir)) %>%
  #mutate(ir = factor(ir, levels = c("Island", "Shore", "Shelf", "Sea"))) %>%
  #ggplot() + geom_density(aes(x=score,y=..scaled..,color=ir)) + theme_classic() + scale_x_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3)) + ylab("Normalized Density") + theme(aspect.ratio=1/4)
  mutate(ir = factor(ir, levels = c("Sea", "Shelf", "Shore", "Island"))) %>%
  ggplot() + geom_violin(aes(y=ir,x=score,fill=ir),draw_quantiles = 0.5) + theme_classic() + scale_x_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3)) + theme(aspect.ratio=1/4,legend.position = "none")
# ggsave("CpGscore_islands_violin.svg")

# plot functional region score distributions
temp <- prepped_data %>% select(Name, UCSC_RefGene_Group, score_median) %>%
  separate_rows(UCSC_RefGene_Group, sep=";") %>%
  distinct() %>%
  dplyr::rename(fr = UCSC_RefGene_Group, score=score_median) %>%
  mutate(fr = case_when(fr=="" ~ "NA",
                        str_detect(fr, "5URT") ~ "5'UTR",
                        str_detect(fr, "3UTR") ~ "3'UTR",
                        TRUE ~ fr)) %>%
  distinct()
keepregions = c("TSS200","5'UTR","Body","3'UTR")
# temp %>%
#   filter(fr %in% keepregions) %>%
#   ggplot() + geom_density(aes(x=score,color=fr)) + theme_classic() + scale_x_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3)) + ylab("Normalized Density")
keepregions = c("3'UTR","Body","5'UTR","TSS200")
temp %>%
  mutate(fr = factor(fr, levels = keepregions)) %>%
  filter(fr %in% keepregions) %>%
  ggplot() + geom_violin(aes(y=fr,x=score,fill=fr), draw_quantiles = 0.5) + theme_classic() + scale_x_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3)) + theme(aspect.ratio=1/4,legend.position = "none")
# ggsave("CpGscore_regions_violin.svg")
rm(temp)
```

@@

# plot: n_sites, mu, betaT, sigmaP, sigmaPT, sigmaT
n_cutlo = 5
n_cuthi = quantile(gene_summary$n, 0.99)
pN <- post_histogram(gene_summary$n,1,c(0,200),paste0("N sites (max = ",max(gene_summary$n),")")) + geom_vline(xintercept=n_cutlo-0.5,linetype = "dashed",color="red") + annotate(geom="text",x=80,y=0.03,size=2,label=paste0("# genes <",n_cutlo," sites:\n", sum(gene_summary$n<n_cutlo)," / ",nrow(gene_summary))) +
geom_vline(xintercept=n_cuthi,linetype = "dashed",color="red") +annotate(geom="text",x=120,y=0.013,size=2,label=sprintf("# genes >%0.1f sites:\n %i / %i (99th pctile)", n_cuthi,sum(gene_summary$n>n_cuthi),nrow(gene_summary)))
gene_summary <- gene_summary %>% filter(n>=n_cutlo) %>% filter(n<n_cuthi)
pM <- post_histogram(gene_summary$mu,0.1,c(-5,5),"Mu")
pB <- post_histogram(gene_summary$betaT,0.1,c(-5,5),"BetaT")
pP <- post_histogram(gene_summary$sigmaP,0.05,c(0,2),"SigmaP")
pPT <- post_histogram(gene_summary$sigmaPT,0.05,c(0,2),"SigmaPT")
pT <- post_histogram(gene_summary$sigmaT,0.05,c(0,2),"SigmaT")
pP1 <- post_histogram(gene_summary$score_median,0.05,c(-3,3),"Score_median")

# arrange each parameter as a subplot
p_all <- ggarrange(pN,pM,pB,pP,pPT,pT,pP1,hjust=0,nrow=3,ncol=3)
print(p_all)
rm(pN,pM,pB,pP,pPT,pT,pP1)

@@

pP1 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_median), alpha=0.01) + ggtitle("Score_median vs N-sites")

# arrange each parameter as a subplot
p_all <- ggarrange(pM,pB,pE,pP,pPT,pT,pP1,hjust=0,nrow=3,ncol=3)
print(p_all)
rm(pM,pB,pE,pP,pPT,pT,pP1)
```

@@

distance_length <- distanced_data %>%
  filter(!is.na(UCSC_RefGene_Name)) %>%
  select(CHR, UCSC_RefGene_Name, cpgdist, c(sigmaP:sigmaT,score_median)) %>%
  mutate(location_bin = cut(cpgdist, c(seq(-1500, 20000, by = 100)), include.lowest = TRUE)) %>%
  group_by(location_bin) %>%
  summarise(mean_sigmaP = mean(sigmaP), mean_sigmaPT = mean(sigmaPT), mean_sigmaT = mean(sigmaT),  mean_score_median = mean(score_median), n_obs = n(), .groups='drop') %>%
  mutate(location_bin = as.numeric(str_extract(location_bin, "(?<=,)(.*)(?=])")))

@@

# again for conservation scores
distance_length %>%
  drop_na() %>%
  pivot_longer(mean_score_median) %>% 
  #mutate(name = factor(name, c("mean_score_median"), c("Score_median"))) %>%
  sample_frac() %>%
  ggplot(aes(location_bin, value, color = name, fill = name)) +
  geom_point(key_glyph = draw_key_rect) +
  ggforce::facet_zoom(location_bin < 5000) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(title = NULL,
       fill = NULL,
       color = NULL,
       x = "Distance from gene start site (binned to 100 bp)",
       y = "Average Conservation Score", 
       caption = "Figure 1: Average conservation scores of single CpGs as a function of position relative to gene."
       ) +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
cat("\n")
# ggsave("CpGdist.svg")

@@

cat("***")
permed_data %>%
  pivot_longer(dplyr::matches("1|3")) %>%
  separate(name, into = c("type", "method"), sep = "_v") %>%
  mutate(method = factor(method, c("1", "3"), c("Naive bootstrap", "Adjusted bootstrap")),
         type = factor(type, c("score_median"), c("Score_median"))) %>%
  drop_na() %>%
  ggplot(aes(value, fill = type)) +
  geom_histogram(bins = 50, color = "grey30") +
  facet_grid(method~type, scales = "free") +
  theme_minimal() +
  labs(x = "p-value",
       y = "Number of genes",
       title = NULL,
       caption = "Figure 2B: Distribution of boot-strapping p-values for genes as a function of conservation score and bootstrap method.")  +
  theme(legend.position = "none") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
```

### 2020-11-24 conserved_genes_and_paths.Rmd
Edits to figures

```{r load, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pander)
library(gt)
library(dplyr)
# set working directory
setwd("/Users/kevinmurgas/Documents/Data+ project/EPIC data")
library(here)

@@

permed_data <- read_csv(here("data-stan_rg3ps3", "permed_data.csv"))
#enrichedGenes <- read_csv(here("prob1genes.csv"))

@@

# Gene Set Enrichment
tidy_gsea <- function(genes) {
  AnnotationDbi::mapIds(org.Hs.eg.db::org.Hs.eg.db, genes, 'ENTREZID', 'SYMBOL') %>%
    ReactomePA::enrichPathway()
}

cat("Begin reactome GSEA...")
ptm<-proc.time()
enrich_reactome <-enrichedGenes %>%
  pull(Gene) %>% tidy_gsea()
proc.time()-ptm

>>>>

```{r load, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pander)
library(gt)
# set working directory
setwd("/Users/kevinmurgas/Documents/Data+ project/EPIC data")
library(here)

@@

permed_data <- read_csv(here("data_final_rg3", "permed_data.csv"))
#enrichedGenes <- read_csv(here("prob1genes.csv"))

@@

# Gene Set Enrichment with Reactome
tidy_reactome <- function(genes) {
  AnnotationDbi::mapIds(org.Hs.eg.db::org.Hs.eg.db, genes, 'ENTREZID', 'SYMBOL') %>%
    ReactomePA::enrichPathway()
}
# GSEA with KEGG
tidy_kegg <- function(genes) {
  AnnotationDbi::mapIds(org.Hs.eg.db::org.Hs.eg.db, genes, 'ENTREZID', 'SYMBOL') %>%
    clusterProfiler::enrichKEGG()
}

cat("Begin reactome GSEA...")
ptm<-proc.time()
enrich_reactome <-enrichedGenes %>%
  pull(Gene) %>% tidy_reactome()
proc.time()-ptm
cat("Begin KEGG GSEA...")
ptm<-proc.time()
enrich_kegg <-enrichedGenes %>%
  pull(Gene) %>% tidy_kegg()
proc.time()-ptm
